---
title: "Getting Started with enoeR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with enoeR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE   # examples require internet; set TRUE locally
)
library(enoeR)
```

## Overview

`enoeR` provides a one-line interface to Mexico's **ENOE** (*Encuesta
Nacional de Ocupación y Empleo*) microdata published by INEGI.  All
editions from 2005 onwards are covered, including the **ENOE-N**
post-COVID redesign (2020 Q3 – 2022 Q4).

---

## 1. Load a single quarter

```{r single}
# All five tables — Q1 2024
q1_2024 <- enoe_load(2024, 1)

# Inspect structure
names(q1_2024)         # "sdem" "coe1t" "coe2t" "hog" "viv"
nrow(q1_2024$sdem)

# Only SDEM and COE1T (faster: same ZIP, fewer extractions)
q1_2024 <- enoe_load(2024, 1, tables = c("sdem", "coe1t"))
```

---

## 2. Batch download: multiple quarters (In Progress)

```{r batch}
# All quarters 2022–2024
panel <- enoe_list(2022, 2024)

# Single year, Q2–Q4 only
panel_2023 <- enoe_list(2023, 2023, start_q = 2, end_q = 4)

# Check metadata
enoe_meta(panel)
```

> **Note:** 2020 Q2 is automatically excluded — INEGI suspended ENOE
> that quarter due to COVID-19 and conducted the telephone ETOE survey
> instead.

---

## 3. Access a specific quarter (In Progress)

```{r access}
# Panel element names follow the pattern "enoe_{year}_t{quarter}"
sdem_2023q1 <- panel$enoe_2023_t1$sdem
```

---

## 4. Extract one table from all quarters  (In Progress)

```{r extract}
sdem_list <- enoe_extract(panel, "sdem")
length(sdem_list)         # one element per successfully downloaded quarter
names(sdem_list)
```

---

## 5. Stack all quarters into a single data.table  (In Progress)

```{r stack}
library(data.table)

sdem_all <- enoe_stack(panel, "sdem")

# Row counts per quarter
sdem_all[, .N, by = period]

# Add derived variables across the whole panel
sdem_all[, log_wage := log(ingocup + 1)]
```
---

## Table reference

| Code   | Module                                          | Key variables                        |
|--------|-------------------------------------------------|--------------------------------------|
| `sdem` | Socio-demographic characteristics               | Age, sex, education, wages, location |
| `coe1t`| Occupation and Employment Questionnaire Part 1  | Employment status, hours worked      |
| `coe2t`| Occupation and Employment Questionnaire Part 2  | Secondary job, earnings detail       |
| `hog`  | Household questionnaire                         | Household composition, size          |
| `viv`  | Dwelling questionnaire                          | Housing characteristics              |

---

## Survey editions and URL patterns

| Period                   | Edition  |
|--------------------------|----------|
| 2005 Q1 – 2020 Q1        | ENOE     |
| 2020 Q2                  | ETOE ⚠  |
| 2020 Q3 – 2022 Q4        | ENOE-N   |
| 2023 Q1 onwards          | ENOE     |
